\vfill World of Warcraft's New Way to Get You Hooked

### 2013-02-13 paper.tex
Abstract is a little better now.
Analysis started on 1.02

% [5%] Abstract (75 Words Max)
% __ One or two sentences of context about your issue.
% __ Provide your focus question.
% __ Mention one or two relevant arguments.
% __ Briefly explain your conclusion and why it is proper ethically. (Optimal total is 7 sentences or so….)

\vfill  %in combination with \newpage this forces the abstract to the bottom of the page
\begin{abstract}
World of Warcraft uses a ``Starter Edition'' which allows players access for free indefinitely with restrictions. \cite{starterEditionNewsRelease} Studies show a correlation between time playing and addiction to Massively Multiplayer Online Role-Playing Games (MMORPGs) like World of Warcraft. \cite{ProblemUsageAmongHighlyEngagedPlayers}\cite{JustAComputerGameRight} Is it ethical for World of Warcraft to use this model?

@@

Blizzard's World of Warcraft is a software system that runs on computers. \cite{WoWSystemReqs} Software systems are written by Software Engineers. \cite{ACMApprovesSECode} According to the Computer Society and the ACM, all Software Engineers should abide by the rules outlined in the the SE Code of Ethics. \cite{ACMApprovesSECode}\cite{SECode} Therefore, Blizzard's World of Warcraft Software Engineers are subject to "the Code." \cite{SECode}

\subsection{Definition of ``addiction''}
Because this paper will often reference psychological addiction, it is important to make sure this notion is concrete before continuing.
A person who is addicted will pathologically pursue rewards and/or relief through substance abuse or other behaviours, in this case playing World of Warcraft. \cite{DefinitionOfAddiction}

``Addiction is characterized by inability to consistently abstain, impairment in behavioral control, craving, diminished recognition of significant problems with one’s behaviors and interpersonal relationships, and a dysfunctional emotional response.'' \cite{DefinitionOfAddiction}

\subsection{SE Code Section 1.02\\Moderate the interests of the software engineer, the employer, the client and the users with the public good.}
Potential to argue that opponents of my position would say that the game leads to addiction and more time spent playing it means higher chances of addiction. Addiction being bad for the ``public good'' and \emph{not} in the ``interests of the client.'' I will respond with sources saying that time does not correlate to addiction and that high levels of play time do not mean addiction.
\subsection{SE Code Section 2.07\\Identify, document, and report significant issues of social concern, of which they are aware, in software or related documents, to the employer or the client.}
Opponents have suggested warning labels and/or contacting players who exhibit signs of addiction. I may have to concede this point, I see nothing wrong with a few added measures.
``Tip: Remember to take all things in moderation (even World of Warcraft!)'' - In-Game Tip! WoW is at least putting some effort forth.
\subsection{SE Code Section 2.09\\Promote no interest adverse to their employer or client, unless a higher ethical concern is being compromised; in that case, inform the employer or another appropriate authority of the ethical concern.}
The opposition will say that promoting more sales while increasing chances for game addiction is adverse to the client. Again I will argue that just because players are hooked and spending money, does not mean they are addicted and their lives are suffering.
\subsection{SE Code Section 4.04\\Not engage in deceptive financial practices such as bribery, double billing, or other improper financial practices.}
Getting young people hooked on games could be seen as deceptive financial practices in that young people lack impulse control and will be more willing to spend their money frivolously. I will argue that a person's money is their own business and it is not Blizzard's responsibility to ensure a consumer is spending their money wisely.

>>>>

% [5%] Abstract (75-100 Words Max)
% __ One or two sentences of context about your issue.
% __ Provide your focus question.
% __ Mention one or two relevant arguments.
% __ Briefly explain your conclusion and why it is proper ethically. (Optimal total is 7 sentences or so….)

\vfill  %in combination with \newpage this forces the abstract to the bottom of the page
\begin{abstract}
Blizzard Entertainment's World of Warcraft has a free ``Starter Edition'' which allows players to access the game indefinitely with restrictions. \cite{starterEditionNewsRelease} Studies show a correlation between time playing and addiction to Massively Multiplayer Online Role-Playing Games (MMORPGs) like World of Warcraft. \cite{ProblemUsageAmongHighlyEngagedPlayers}\cite{JustAComputerGameRight} Is it ethical for World of Warcraft to use this model?

@@

To aid in my analysis of whether or not Blizzard Entertainment's World of Warcraft trial model is ethical, the Software Engineering Code of Ethics (SE Code) will be used. The SE Code applies to ``professional software engineers, including practitioners, educators, managers, supervisors and policy makers, as well as trainees and students of the profession.'' \cite{SECode} Blizzard Entertainment's World of Warcraft is a software system that runs on computers. \cite{WoWSystemReqs} Software systems are written by software engineers. \cite{ACMApprovesSECode} According to the Computer Society and the Association for Computing Machinery (ACM), all professional software engineers should abide by the rules outlined in the the SE Code. \cite{ACMApprovesSECode}\cite{SECode} Blizzard Entertainment can be seen as a ``practitioner'' in addition to some of its employees who are professional software engineers. Therefore, Blizzard Entertainment and Blizzard Entertainment's software engineers are subject to the SE Code. \cite{SECode} Being subject to the SE Code allows me to scrutinize their actions under it.

\subsection{Definition of ``addiction''}
Because this paper will often reference psychological addiction, it is important to make sure this notion is concrete before continuing.
A person who is addicted will pathologically pursue rewards and/or relief through substance abuse or other behaviors, in this case playing World of Warcraft. \cite{DefinitionOfAddiction}

Addiction can be seen as an inability to abstain from something, impairment in behavior control regarding this thing, and a craving for this thing. There will be a diminished recognition of issues with one's behavior, suffering of interpersonal relationships, and a dysfunctional emotional response. \cite{DefinitionOfAddiction}

\subsection{SE Code Section 1.02\\Moderate the interests of the software engineer, the employer, the client and the users with the public good.}
In the next few sections this tenant is broken down into it's parts. I will explore the ``interests of the software engineer,'' in this case World of Warcraft's software engineers, the ``interests of ... the employer,'' in this case Blizzard Entertainment, and lastly the ``interests of ... the client and the users,'' the player. Following that will be a discussion of the ``public good'' using Utilitarianism and the benefits of videogames to guide my definition.
\subsubsection{Interests of the Software Engineer}
MORE INFO HERE
\subsubsection{Interests of Blizzard Entertainment}
Blizzard Entertainment is a publicly traded company. \cite{BlizzStock} As such its main goal is to give its shareholders profits and increase the volume of business it handles. In no uncertain terms, Blizzard Entertainment wants as many people to buy and play World of Warcraft (and its other products) as it can get. Since World of Warcraft is subscription-based \cite{WoWSubscription}, Blizzard Entertainment sees money every single month from World of Warcraft players. None of the other games that Blizzard makes are subscription based \cite{BlizzGameList} and as such World of Warcraft makes Blizzard Entertainment the most money of them all due to its recurring nature. Blizzard Entertainment's interests are in getting more players to buy their games, particularly World of Warcraft.
\subsubsection{Interests of the Player}
By purchasing a game like World of Warcraft, players are committing themselves to pay Blizzard Entertainment every month. Some players are reaping benefits that they would not be able to get otherwise, such as social interactions. \cite{IsThereEvidenceOfInternetAddiction}
MORE INFO HERE
\subsubsection{The ``Public Good''}
In order to define the ``public good'' I assert that Utilitarianism is the most effective ethical system to use. Utilitarianism states that ``[t]he creed which accepts as the foundation of morals, Utility, or the Greatest Happiness Principle, holds that actions are right in proportion as they tend to promote happiness, wrong as they tend to produce the reverse of happiness.'' \cite{Utilitarianism}
MOER INFO HERE
\subsubsection{Ethical Conclusion}
MORE INFO HERE

Potential to argue that opponents of my position would say that the game leads to addiction and more time spent playing it means higher chances of addiction. Addiction being bad for the ``public good'' and \emph{not} in the ``interests of the client.'' I will respond with sources saying that time does not correlate to addiction and that high levels of play time do not mean addiction.

\subsection{SE Code Section 2.07\\Identify, document, and report significant issues of social concern, of which they are aware, in software or related documents, to the employer or the client.}
Opponents have suggested warning labels and/or contacting players who exhibit signs of addiction. I may have to concede this point, I see nothing wrong with a few added measures.
``Tip: Remember to take all things in moderation (even World of Warcraft!)'' - In-Game Tip! WoW is at least putting some effort forth.

\subsection{SE Code Section 2.09\\Promote no interest adverse to their employer or client, unless a higher ethical concern is being compromised; in that case, inform the employer or another appropriate authority of the ethical concern.}
The opposition will say that promoting more sales while increasing chances for game addiction is adverse to the client. Again I will argue that just because players are hooked and spending money, does not mean they are addicted and their lives are suffering.

\subsection{SE Code Section 4.04\\Not engage in deceptive financial practices such as bribery, double billing, or other improper financial practices.}
Getting young people hooked on games could be seen as deceptive financial practices in that young people lack impulse control and will be more willing to spend their money frivolously. I will argue that a person's money is their own business and it is not Blizzard's responsibility to ensure a consumer is spending their money wisely.
